---
title: "Different Random Forest Packages in R"
author: "Thiyanga Talagala"
date: "14 November 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(knitr)
opts_chunk$set(tidy = TRUE)
```

# randomForest Package

## Data pre-processing

### Split iris data to Training data and testing data

```{r}
ind <- sample(2,nrow(iris),replace=TRUE,prob=c(0.7,0.3))
trainData <- iris[ind==1,]
testData <- iris[ind==2,]

```

1. Load randomForest

```{r}
library(randomForest)

```

2. Generate Random Forest learning treee

```{r}
iris_rf <- randomForest(Species~.,data=trainData,ntree=100,proximity=TRUE)
table(predict(iris_rf),trainData$Species)

```

3. Try to print Random Forest model and see the importance features

```{r}
print(iris_rf)

plot(iris_rf)
```

4. Assessing model fit

```{r}
importance(iris_rf)
varImpPlot(iris_rf)

```

5. Predict the class labels for test data

```{r}
irisPred<-predict(iris_rf,newdata=testData)
irisPred
table(irisPred, testData$Species)

```

6. Try to see the margin, positive or negative, if positif it means correct classification

```{r}
plot(margin(iris_rf,testData$Species))

```

7. Tune randomForest for the optimal mtry parameter

method 1

```{r}
tune.rf <- tuneRF(iris[,-5],iris[,5], stepFactor=0.5)
print(tune.rf)

```

method 2

We can also tune the structure, ie, finding the best hyperparameters of the method via grid search

```{r}
library(e1071)
tuned.r <- tune(randomForest, train.x=Species~., data=trainData, validation.x=testData)
best.model <- tuned.r$best.model
best.model
predictions <- predict(best.model, testData)
table.random.forest <- table(testData$Species, predictions)
table.random.forest
```




8. Tree size

```{r}
hist(treesize(iris_rf))
```

# The randomForestSRC Package

```{r}
library(randomForestSRC)
```

1. Fitting a random forest

```{r}
## Edgar Anderson's iris data
iris.obj <- rfsrc(Species ~., data = trainData)
iris.obj
plot(iris.obj)

# obtain class labels


```

2. predict based on the results of rfsrc

```{r}

rfsrcpred <- predict(iris.obj, testData)

predictions<- rfsrcpred$predicted
predictions


```

3. Plotting commands

```{r}

data(iris.obj, package="ggRandomForests")
gg_dta<- gg_rfsrc(iris.obj)

plot(gg_dta)
```




## Parallel execution with random forest

```{r}
library(randomForest)
library(doParallel)
workers <- detectCores() 
workers
cl <- makePSOCKcluster(workers)
registerDoParallel(cl)

x <- matrix(runif(500), 100)
y <- gl(2, 50)
ntree <- 1000

rf <- foreach(n=rep(ceiling(ntree/workers), workers),
              .combine=combine, .multicombine=TRUE,
              .packages='randomForest') %dopar% {
  randomForest(x, y, ntree=n)
}

```




